---
title: "Let's Go To The Movies"
author: 'Davin Dillon'
date: "2022-05-03"
output:
  slidy_presentation:
    font_adjustment: 1
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

```{r, include = F}

library(tidyverse) # multiple tidy packages
library(readr) # for reading files
library(ggplot2) # for plotting
library(plotly) # for interactive plotting
library(dplyr) # data manipulation
library(flextable) # pretty tables
library(formattable) # comma function formatting numbers
library(ggforce) # facet columns
library(gridExtra) # side by side plots

```



```{r, include = F}

budget <- read_csv('movies_budget.csv') # budget including new movies
oscars <- read_csv('the_oscar_award.csv') # oscar winners and losers

meta <- read_csv('movies_metadata.csv') # majority of data used

inflation <- read_csv('inflation_data.csv') 
# inflation data for calculations

```

<span style = "color:navy;, font-family:Arial, Helvetica, sans-serif;"><h2>Paying for Oscars or Oscars for Pay?</h2><ul><li>The largest part of my data comes from a data set I got from github. [<a style = 'color:green;'>yash91sharma github</a>](https://raw.githubusercontent.com/yash91sharma/IMDB-Movie-Dataset-Analysis/master/movie_metadata.csv) created the data for a project similar to mine in which they asked which countries made the most movies.This data set contains `r comma(nrow(meta), digits = 0)` rows of data and seems to have been made around 2017.</li>

<li>I used [<a style = 'color:green;'>alexsychu kaggle</a>](https://www.kaggle.com/code/alexsychu/movie-industry-eda-alex-chu-5-april-2022/data) to gather some more data up to 2020. This second dat set had an additional `r comma(nrow(budget), digits = 0)` rows of data. This user was trying to answer the question of how to predict if a movie will do well.

</li><li>Lastly for movie data, I used data from [<a style = 'color:green;'>unanimad kaggle</a>](https://www.kaggle.com/datasets/unanimad/the-oscar-award) . This data set had `r comma(nrow(oscars),digits = 0)` rows of data and this kaggle user asked various questions about who won Oscars.

</li><li>My last and only 'non-movie' data set comes from [<a style = 'color:green;'>officialdata.org</a>](https://www.officialdata.org/us/inflation/1800?amount=1#:~:text=Value%20of%20%241%20from%201800%20to%202022&text=The%20dollar%20had%20an%20average,Labor%20Statistics%20consumer%20price%20index.) , where I obtained 223 rows of inflation data which allowed me to calculate budgets and gross revenue with inflation as a factor.</li></ul></span>


---

<span style = "color:navy;, font-family:Arial, Helvetica, sans-serif;"><h2>Data Wrangling</h2>
There were many renames and some work to choose a movie budget in some cases. Once I had the columns I wanted, I set in on joining the data together. First, I joined the meta data with the budget data to include all of the titles for which I had information. My next adventure (or misadventure) was to join this data with the inflation data. Once that was done, all that was left was joining this information with the Oscar data. All in all, I used three full joins. </span>



```{r, echo = T, warning = F}

oscars <- oscars %>%    # rename year for join
  rename('year' = 'year_film')

meta$year <-  format(as.Date(meta$release_date, format = "%m/%d/%Y"), "%Y")

blue = '#000080'
  
meta <- meta %>% 
  mutate(year = as.numeric(year))

# renames
budget <- budget %>% 
  rename(vote_average = score) %>% # renaming vote data for joins etc
  rename(vote_count = votes) %>% 
  rename(Title = 'Movie Title')


oscars <- oscars %>% 
  rename(Title = 'film')  # rename for joins

adj <- inflation %>% 
  mutate(multiplier = (22.82/amount)) 
# create multiplier column for easy calculations

budget <- budget %>% 
  rename(new_budget = Budget)  # rename for joins etc


options(scipen = 100) # avoid scientific notation 

full_budget <- full_join(meta, budget, on = 'Title') 
# full join of metadata and budget data to get new and old movies etc

full_bud <- full_budget %>% 
  mutate(budget = pmax(new_budget, meta_budget, na.rm = T)) %>% 
  select(Title,genres, budget,new_budget, meta_budget, popularity,year,
         release_date, revenue, runtime, vote_average,
         vote_count,gross)
# set budget to max of two different budgets.
# picking max is arbitrary, but needed in most cases


full_bud <- full_bud[-c(1,2,3),] %>%
   arrange(desc(as.numeric(budget)))
# remove first three unnecessary rows


# format(date, format="%Y")


full_bud <- full_bud %>% 
  mutate(year = replace(year, year == 1900, 2022))


adj_bud <- full_join(full_bud, adj,
                             on = c('Title', 'year')) %>% 
  mutate(with_inflation = (as.numeric(budget) * (multiplier))) %>% 
  mutate(gross_inflation = (as.numeric(full_bud$gross)
                            * (multiplier))) %>% 
  select(Title,genres,vote_average, vote_count, budget,
         with_inflation, gross_inflation, year, gross,
         release_date) %>% 
  arrange(desc(with_inflation))  


osc_bud <-  full_join(adj_bud, oscars, on  = c('Title'))


```




```{r, include = F}
knitr::opts_chunk$set(echo = F)
noms <- osc_bud %>% 
  distinct(Title)


osc_pop <-  osc_bud %>% 
  mutate(vote_count = replace(vote_count, vote_count < 10000, NA)) %>%           filter(vote_count != 0) %>% 
  filter(!is.na(vote_count))
# filtered out movies that had less than 10000 votes  
# for popular  movie set 

total_votes <- sum(full_bud$vote_count, na.rm = T)
```




------------------------------------------------------------------------

<span style = "color:navy;, font-family:Arial, Helvetica, sans-serif;"><h2>Some Dataset Numbers</h2></span>

```{r, include = T}
# glimpse(oscar_titles)
# glimpse(oscar_wn_titles)
# glimpse(bst_act %>% distinct(Title))
# glimpse(bst_pic)
# glimpse(bst_pic_won)
# glimpse(noms) 

```


```{r, include = F}

oscar_titles <- oscars %>% # count number of titles in oscar data
  distinct(Title)

oscar_wn_titles <- oscars %>% # count number of oscar winning movies
  filter(winner == T) %>% 
  distinct(Title)

total_noms <- osc_bud %>% 
  filter(!is.na(winner))



total_wins <- osc_bud %>% 
  filter(winner == T)


bst_pic <-  osc_bud %>% 
  group_by(Title) %>% 
  filter((str_detect(category, 'PIC') | 
            str_detect(category, 'OUT'))) %>% 
  select(Title, year,budget, with_inflation,
         gross_inflation, gross, genres, vote_count,
         vote_average, category, winner) %>% 
  distinct(Title, .keep_all = T)


bst_pic_won <- bst_pic %>% 
  filter(winner == T) %>% 
  distinct(Title, .keep_all = T)

bst_act <- osc_bud %>% 
  group_by(Title) %>% 
  filter(str_detect(category, 'ACT')) %>% 
  select(Title, genres, budget, category, winner)

bst_act_wn <- bst_act %>% 
  filter(!is.na(winner)) %>% 
  filter(winner == T) %>% 
  distinct(Title, .keep_all = T)

```


[The data I was able to collect contains `r comma(nrow(oscar_titles), digits = 0)` movies nominated for an Academy Award since its inception in 1929. Of these `r comma(nrow(oscar_titles),digits = 0)` movies. `r comma(nrow(oscar_wn_titles), digits = 0)` won at least one award. There have been `r comma(nrow(total_noms), digits = 0)` total Oscar nominations, and `r comma(nrow(total_wins), digits = 0)` total Oscar wins in the dataset. `r comma(nrow(bst_pic), digits = 0)` movies have been nominated for Best Picture in its many forms. Out of these, `r comma(nrow(bst_pic_won), digits = 0)` won. `r comma(nrow(bst_act %>% distinct(Title)), digits = 0)` movies have had an actor or actress nominated in either a leading or supporting role.  `r comma(nrow(bst_act_wn), digits = 0)` movies had at least one winner in an acting category. ]{style="color:navy;, font-family:Arial, Helvetica, sans-serif;"}


---


<span style = "color:navy;, font-family:Arial, Helvetica, sans-serif;"><h2>Plotting Profit</h2></span>

```{r, include = T}

osc_bud <- osc_bud[!is.na(osc_bud$budget), ]
osc_bud <- osc_bud %>% 
  mutate(budget = as.numeric(budget))
p <- osc_bud %>% select(Title, genres, gross, budget, winner) %>% 
  ggplot(aes(x = budget, y = gross, color = winner, label = Title), alpha = 0.1) +
  geom_abline(intercept = 0, slope = 1) +
  geom_point() +
  theme(axis.text.x = element_text(angle = -90,
                                   size = 5, color = blue ),
        axis.text.y = element_text(color = blue)) +
  theme(axis.title.x = element_text(color = blue),
        axis.title.y = element_text(color = blue))

ggplotly(p)
```


---


<span style = "color:navy;, font-family:Arial, Helvetica, sans-serif;"><h2>Highest Percent Profit</h2></span>

```{r, include = T}
high_prof <- osc_bud %>% 
  filter(budget != 0) %>% 
  mutate(pct_prof = 100 * (as.numeric(gross)/as.numeric(budget))) %>% 
  select(Title, budget, gross, pct_prof) %>% 
  arrange(desc(pct_prof))

flextable(head(high_prof, 50))

```

---

<span style = "color:navy;, font-family:Arial, Helvetica, sans-serif;"><h2>Lowest Percent Profit</h2></span>

```{r}

low_prof <- high_prof %>% 
  na.omit(budget) %>% 
  arrange(pct_prof)

flextable(head(low_prof, 50))

```


---


<span style = "color:navy;, font-family:Arial, Helvetica, sans-serif;"><h2>Who spends more money?</h2></span>



```{r , warning = FALSE}

osc_bud <- osc_bud %>%
  mutate(budget = as.numeric(budget)) %>%
  mutate(gross = as.numeric(gross))

win_prof <- osc_bud %>%
  filter(winner == T) %>%
  filter(!is.na(budget)) %>%
  filter(!is.na(gross)) %>%
  mutate(winner_profit = (100*sum(gross))/(sum((budget)))) %>% 
  mutate(percent = 100 * (gross/budget))

lsr_prof <- osc_bud %>%
  filter(winner == F) %>%
  filter(!is.na(budget)) %>%
  filter(!is.na(gross)) %>%
  mutate(loser_profit = 100*sum(gross/sum(budget))) %>% 
  mutate(percent = 100 *(gross/budget))

no_nom_prof <- osc_bud %>%
  filter(is.na(winner)) %>%
  filter(!is.na(budget)) %>%
  filter(!is.na(gross)) %>%
  mutate(nonnom_profit = (100*sum(gross))/(sum(budget))) %>% 
  mutate(percent = 100 * (gross/budget))


money_plot <- osc_bud %>% select(Title, genres, gross,
                                 budget,with_inflation,
                                 gross_inflation, winner) %>%
  ggplot() +
  geom_density(aes((budget), fill = winner), alpha = 0.4)   +
  theme(axis.text.x= element_text(angle=-90)) +
  xlim(0, 60000000) +
  ylim(0, 0.00000035) +
  theme(axis.text.x = element_text(angle = -90,
                                   size = 5, color = blue ),
        axis.text.y = element_text(color = blue)) +
  theme(axis.title.x = element_text(color = blue),
        axis.title.y = element_text(color = blue))

ggplotly(money_plot)


money_plot2 <- osc_bud %>% select(Title, genres,
                                  gross, with_inflation,
                                  budget, winner, 
                                  gross_inflation) %>%
  ggplot() +
  geom_density(aes(with_inflation, fill = winner), alpha = 0.4)+
  theme(axis.text.x=element_text(angle=-90)) +
  xlim(0, 60000000) +
  ylim(0, 0.00000035) +
  theme(axis.text.x = element_text(angle = -90,
                                   size = 5, color = blue ),
        axis.text.y = element_text(color = blue)) +
  theme(axis.title.x = element_text(color = blue),
        axis.title.y = element_text(color = blue))

ggplotly(money_plot2)

grid.arrange(money_plot, money_plot2, ncol=2)

```



<span style = "color:navy; font-family:Arial, Helvetica, sans-serif;"><ul><li>The average Oscar winner budget was $`r comma(mean(win_prof$with_inflation), digits = 2)`.</li><li>The median Oscar winner budget was $`r comma(median(win_prof$with_inflation), digits = 2)`</li><li>The average Oscar loser budget was $`r comma(mean(lsr_prof$with_inflation), digits = 2)`</li><li>The median Oscar loser budget was $`r comma(median(lsr_prof$with_inflation), digits = 2)`</li><li>The average non-nominated budget was $`r comma(mean(no_nom_prof$with_inflation, digits = 2))`.</li><li>The median non-nominated budget was `r comma(median(no_nom_prof$with_inflation), digits = 2)`</li></ul></span>


---

<span style = "color: navy;, font-family:Arial, Helvetica, sans-serif;"><h2>Who makes more money?</h2></span>

```{r, warning = F}

prof_plot <- osc_bud %>% select(Title, genres, gross,
                                budget,with_inflation,
                                gross_inflation,  winner) %>% 
  ggplot() +
  geom_density(aes((gross), fill = winner), alpha = 0.4)   + 
  xlim(0, 200000000) + 
  ylim(0, 0.000000070) +
  theme(axis.text.x = element_text(angle = -90,
                                   size = 5, color = blue ),
        axis.text.y = element_text(color = blue)) +
  theme(axis.title.x = element_text(color = blue),
        axis.title.y = element_text(color = blue))

ggplotly(prof_plot)


prof_plot2 <- osc_bud %>% select(Title, genres, gross,
                                budget,gross_inflation, winner) %>% 
  ggplot() +
  geom_density(aes(gross_inflation, fill = winner), alpha = 0.4)   +     xlim(0, 200000000) +
  ylim(0, 0.000000070) +
  theme(axis.text.x = element_text(angle = -90,
                                   size = 5, color = blue ),
        axis.text.y = element_text(color = blue)) +
  theme(axis.title.x = element_text(color = blue),
        axis.title.y = element_text(color = blue))


ggplotly(prof_plot2)

grid.arrange(prof_plot, prof_plot2, ncol=2)


```

<span style = "color:navy; font-family:Arial, Helvetica, sans-serif;"><ul><li>The average Oscar winner percent profit was `r comma(win_prof$winner_profit[1], digits = 2)`.</li><li>The median percent profit for Oscar winners was `r comma(median(win_prof$percent),digits = 2)`</li><li>The average Oscar loser percent profit was `r comma(lsr_prof$loser_profit[1], digits = 2)`. </li><li>The median percent profit for Oscar losers was `r comma(median(lsr_prof$percent), digits = 2)`. </li><li>The average non-nominated percent profit was `r comma(no_nom_prof$nonnom_profit[1], digits = 2)`.</li><li>The median percent profit for non-nominated movies was `r comma(median(no_nom_prof$percent),digits = 2)`</li></ul></span>


---

<span style = "color:navy; font-family:Arial, Helvetica, sans-serif;">
<h2> Spent vs Made </h2></span>



```{r, warning = F}

grid.arrange(money_plot, prof_plot, ncol=2)

grid.arrange(money_plot2, prof_plot2, ncol=2)


```

---


<span style = "color:navy; font-family:Arial, Helvetica, sans-serif;"><h2>Best Picture Budgets with Inflation</h2></span>



```{r}

bst_pic_budgets <- bst_pic_won %>% 
  filter(budget > 0) %>% 
  mutate(year_title = paste(year, ', ', Title))

r <- bst_pic_budgets %>%  
  ggplot() +
  geom_col(aes(x = year_title, y = with_inflation), fill= blue) +
  theme(axis.text.x = element_text(angle = -90,
                                   size = 5, color = blue ),
        axis.text.y = element_text(color = blue)) +
  theme(axis.title.x = element_text(color = blue),
        axis.title.y = element_text(color = blue))

ggplotly(r)

min_bud <- bst_pic_won %>% 
  filter(with_inflation > 0) %>% 
  filter(with_inflation == min(with_inflation)) %>% 
  select(Title, with_inflation) %>% 
  arrange(with_inflation)

min_bud_title <- head(min_bud$Title, 1)
min_bud_cost <- head(min_bud$with_inflation, 1)

avg_bst <- mean(min_bud$with_inflation)

max_bud <- bst_pic_won %>% 
  filter(with_inflation == max(with_inflation)) %>% 
  select(Title, with_inflation) %>% 
  arrange(desc(with_inflation))

max_bud_title <- head(max_bud$Title, 1)
max_bud_cost <- head(max_bud$with_inflation, 1)


```



The minimum budget for a Best Picture winner with inflation was `r min_bud_title` with a cost of $`r comma(min_bud_cost, digits = 2)`. The average budget for a Best Picture winner with inflation was $`r comma(avg_bst, digits = 2)`. The maximum budget for a Best Picture winner with inflation was `r max_bud_title` with a cost of $`r comma(max_bud_cost, digits = 2)`.



